---
title: "Short-term associations of road density and road features with in-vehicle PM2.5 during
  daily trips"
subtitle: Main text figures
output:
  pdf_document:
    toc: yes
    keep_tex: true
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = F)
```

```{r}
# libraries
library(tidyverse)
library(knitr)
library(here)
library(ggthemes)
library(DescTools)
library(RColorBrewer)
library(forcats)
library(lubridate)
library(patchwork)

figloc <- "~/Dropbox/CommuteVar/figures"

# from roadiness-eda.Rmd
load(here("figures/eda-tabfig.RData"))
# from figure2.R
load(here("figures/fighourly.RData"))

# paired analysis
load(file.path(figloc, "figpaired.RData"))

# masstime
load(here("results/masstime.RData"))


# lm res

load(here("results/lmfig.RData"))


# reject ER changes
load(here("results/er-reject-res.RData"))

# quantiles
load( here("results/sens-er.RData"))

# maps
load(here("data/maps.RData"))
```




# Tables

```{r}
k <- 1

```



## Table `r k`. Correlations between road features including standardized road density (for highways and local roads at 1 km and 500 m resolution) and speed (MPH).

```{r}
k <- k + 1
```

```{r}
options(knitr.kable.NA = '')
corr1c <- corr1b[c(1 : 4), c(1 : 4)]
corr1c <- data.frame(corr1c)
corr1c <- mutate(corr1c, Variable = "Road density", Road = rep(c("Highway", "Local"), 2),
                 Resolution = rep(c("1 km", "500 m"), each = 2)) %>% 
  dplyr::select(Variable, Road, Resolution, everything())
rownames(corr1c) <- NULL

kable(corr1c, digits = 2)
```

## Table `r k`. Unadjusted differences in in-vehicle PM$_{2.5}$ (log $\mu$g/m$^3$) comparing each quartile of road density (Q2, Q3, Q4) to the lowest quartile for highways and local roads at 1 km and 500 m resolution from linear mixed models.

```{r}
k <- k + 1
```

```{r}
kable(arrange(qsrnesstab, Resolution, Road))

```


# Figures
```{r}
k <- 1

```


## Figure `r k`. Map of standardized road density in the Washington, DC metropolitan area for A. highways and B. local roads at 1 km resolution.

```{r}
k <- k + 1
```


```{r}
hw1 + ggtitle("A. Road density: Highway") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12), title = element_text(size = 10)) +
  loc1 + ggtitle("B. Road density: Local") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12), title = element_text(size = 10))

ggsave(file = here("figures/fig1-v2.png"), width = 8, height = 3.5)

```

## Figure `r k`. Variation in road density for highways and local roads at 1 km resolution (in standard deviations (SD)) and over each observed trip (N=69).  

```{r}
k <- k + 1
```


```{r}
rcomm1 <- tripplot[[2]]
```

```{r}
srness1 <- dplyr::select(rcomm1, ID, id2, timemin, contains("srness")) %>%
  pivot_longer(-c(ID, id2, timemin)) %>%
  dplyr::filter(name %in% c("srness1loc", "srness1hw")) %>%
  mutate(name = case_when(name == "srness1loc" ~ "Road density: Local",
                          name == "srness1hw" ~ "Road density: Highway")) 



srness <- ggplot(srness1, aes(x = timemin, y = id2, fill = value)) +
  geom_tile() + ylab("Trip") + xlab("Minutes since trip start") +
 scale_fill_gradient2(name = "Density (SD)", low = "#dedc59", high = "#5c3a85", mid = "white") +
  facet_wrap(~name) +
  theme(text = element_text(size=12), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position  = c(.9, .3))  + theme(axis.text.y = element_blank(),
                                        axis.ticks.y = element_blank(),
                                        legend.title = element_text(size = 10))

srness
ggsave(file = here("figures/fig2-v2.png"), width = 5.5, height = 3.5)

```



## Figure `r k`. Proportion of road type by quartiles of road density (for highways and local roads at 1 km and 500 m resolution).

```{r}
k <- k + 1
```


```{r}
rtypecorr 

ggsave(file = here("figures/fig3-v2.png"), width = 6, height = 3.5)
```


## Figure `r k`. Associations of road density with in-vehicle PM$_{2.5}$ concentrations from linear mixed models reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of road density (Q2, Q3, Q4) to the lowest quartile (Q1) for highways and local roads at 1 km resolution. Results include main (unadjusted) models as well as models that separately adjusted for meteorology, ambient PM$_{2.5}$, rush hour trips, and all road features.

```{r}
k <- k + 1

```


```{r}
q1km

ggsave(file = here("figures/fig4-v2.png"), width = 6, height = 3.5)

```


## Figure `r k`. Associations of road density with in-vehicle PM$_{2.5}$ concentrations from linear mixed models for main models (lag 0) and lags up to 10 minutes, reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of road density (Q2, Q3, Q4) to the lowest quartile (Q1) for highways and local roads at 1 km resolution.


```{r}
k <- k + 1
```


```{r}
roadlag1hw 

ggsave(file = here("figures/fig5-v2.png"), width = 6, height = 6)

```


