---
title: 'Short-term associations of road density and road features with in-vehicle PM2.5 during daily trips'
subtitle: "Supplementary Material"
output: 
  pdf_document:
    toc: true
    keep_tex: true
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = F)
```

```{r}
# libraries
library(tidyverse)
library(knitr)
library(here)
library(viridis)
library(ggthemes)
library(DescTools)
library(RColorBrewer)
library(forcats)
library(lubridate)
library(patchwork)

source(here("rcode/windrose.R"))

# from roadiness-eda.Rmd
load(here("figures/eda-tabfig.RData"))
# from figure2.R
load(here("figures/fighourly.RData"))

# supplement from roadiness-eda.Rmd
load(here("results/supp-plots.RData"))


# gps plot data
load(here("results/gpsplotdata.RData"))

# masstime: NOTE THIS IS NOT UPDATED
# load(here("results/masstime.RData"))

# lag figs
load(here("results/lmfig.RData"))


# icc : NOT UPDATED/NEEDED
# load(here("results/icc.RData"))


# reject ER changes
load(here("results/er-reject-res.RData"))


# quantiles
load( here("results/sens-er.RData"))

```


```{r, cache = T}
# maps
load(here("data/maps.RData"))
```

# Tables

```{r}
k <- 1
```

## Table S`r k`. Demographic characteristics (N (%)) for N=25 participants.

```{r}
k <- k + 1
```


```{r}
kable(demo)
```




## Table S`r k`. Meteorological characteristics (N (%)) for N=45 person-days of observation.

```{r}
k <- k + 1
```

```{r}
kable(catweath)
```

\blandscape

## Table S`r k`. Correlations of road features with ambient PM$_{2.5}$ and daily meteorology. Variables include: Road density: road density for highways and local roads at 1 km and 500 m resolution (standard deviations); Speed (MPH); DailyPM: daily ambient PM$_{2.5}$ ($\mu$g/m$^3$);  HourlyDevPM: hourly deviations from daily ambient PM$_{2.5}$ ($\mu$g/m$^3$); WindSpeed (m/s); MaxTemp: maximum temperature ($^{\circ}$C); MinTemp: minimum temperature ($^{\circ}$C); MeanTemp: mean temperature ($^{\circ}$C); and Humid: relative humidity (%)).

```{r}
k <- k + 1
```

```{r}
options(knitr.kable.NA = '')
corr1c <- corr1b[, -c(1 : 4)]
corr1c <- data.frame(corr1c)
corr1c <- mutate(corr1c, Variable = rownames(corr1c),
                 Variable = ifelse(substr(Variable, 1, 2) == "RD", paste0("Road density: ", rep(c("Highway", "Local"), 2), ", ",
                                  c(rep(c("1 km", "500 m"), each = 2))), Variable)) %>%
  dplyr::select(Variable, everything())
rownames(corr1c) <- NULL

kable(corr1c, digits = 2)
```

## Table S`r k`. Quantiles (Q1, Q2, Q3, Q4) for road density (for highways and local roads at 1 km and 500 m resolution) and speed.

```{r}
k <- k + 1
```

```{r}
quant1 <- quanttab[c(2 : 5, 1),]
quant1 <- mutate(quant1, Min = round(Min, 2),
                 P25 = round(P25, 2),
                 P50 = round(P50, 2),
                 P75 = round(P75, 2),
                 Max = round(Max, 2),
                 Q1 = paste0("(", Min, ", ", P25, "]"),
                 Q2 = paste0("(", P25, ", ", P50, "]"),
                 Q3 = paste0("(", P50, ", ", P75, "]"),
                 Q4 = paste0("(", P75, ", ", Max, "]")) %>%
  dplyr::select(-c(Min : Max))
kable(quant1, digits = 2)
```

\elandscape

# Figures

```{r}
k <- 1
```


## Figure S`r k`. Map of standardized road density in the Washington, DC metropolitan area for A. highways and B. local roads at 500 m resolution.

```{r}
k <- k + 1
```


```{r, cache = T}
hw500 + ggtitle("A. Road density: Highway") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12), title = element_text(size = 10)) +
  loc500 + ggtitle("B. Road density: Local") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12), title = element_text(size = 10))

```

## Figure S`r k`. Distributions of in-vehicle PM$_{2.5}$ exposures ($\mu$g/m$^3$) for each participant (N=25) on the natural log scale, along with observation time in minutes.  Participants are ordered by their median in-vehicle PM$_{2.5}$ exposures across trips.

```{r}
k <- k + 1
```

```{r}
rcomm1 <- boxplot1[[2]]
ns <- boxplot1[[3]]

dates <- dplyr::select(rcomm1, date) %>% unique() %>% 
  rename(date_local = date)
rcomm2 <- rcomm1 %>%
  mutate(PM = PM + 0.01)
figs1 <- ggplot(rcomm2, aes(y = fct_reorder(ID, medl), x  = PM)) + geom_boxplot() +#outlier.shape = NA
   ylab(expression(paste("Minutes observed per participant"))) +
  xlab((expression(paste("PM"[2.5], " (", mu, "g/m"^3, ")")))) +
  scale_x_continuous(trans = "log", breaks = c(0.08, 0.5, 3, 20)) +
  theme_bw() +
  scale_y_discrete(breaks = ns$ID, labels = ns$n) +
  theme(text = element_text(size = 12))
figs1 

```



## Figure S`r k`. Percentage of observations on each road type per participant (N=25).

```{r}
k <- k + 1
```

```{r}
perc <- roadtype1[[2]]
perc <- mutate(perc, rtype = factor(rtype, levels = c("High/SecHigh", "Other", "LocalConn", "Local"),
                                    labels = c("Highway", "Ramp/Tunnel",
                                               "LocalConn", "Local"))) 

roadtypep <- ggplot(perc, aes(x= rtype, y = fct_inorder(ID), fill= perc)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red", name = "Percent", breaks = c(0, 25, 50, 75, 100),
                      labels =  paste0(c(0, 25, 50, 75, 100),  "%"),
                      limits = c(0, 100)) +
  xlab("") + ylab("Participant") +
  theme(text = element_text(size=12), axis.text.y = element_blank())

roadtypep

```

## Figure S`r k`. Variation in PM$_{2.5}$ (in standard deviations (SD) from trip-specific means) over each observed trip (N=69).  

```{r}
k <- k + 1
```



```{r, fig.width = 3.5}
rcomm1 <- tripplot[[2]]
pmtile <- ggplot(rcomm1, aes(x = timemin, y = id2, fill = lPM2)) +
  geom_tile() + ylab("Trip") + xlab("Minutes since trip start") +
  # scale_fill_gradient2(name = "Change in PM2.5 (SD)", low = "green", high = "red",
                       # mid = "white") +
  scale_fill_gradient2(name = "Change in PM2.5 (SD)", low = "#dedc59", high = "#5c3a85", mid = "white") +

  #theme_bw() +
  theme(text = element_text(size=12), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), legend.position = c(.75, .25))

pmtile


```

## Figure S`r k`. Variation in A. road type (highway, ramp/tunnel, local connecting, and local roads) and B. speed (mph) over each observed trip (N=69).

```{r}
k <- k + 1
```


```{r}
rcomm1 <- tripplot[[2]]
```

```{r}
roadtype2 <- ggplot(rcomm1, aes(x = timemin, y = id2, fill = rtype2)) +
  geom_tile() + ylab("Trip") + xlab("Minutes since trip start") +
  scale_fill_colorblind(name = "Road type") +
 # theme_bw() +
  theme(text = element_text(size=12), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position  = c(.75, .25))

speed <- ggplot(rcomm1, aes(x = timemin, y = id2, fill = (mph))) +
  geom_tile() + ylab("Trip") + xlab("Minutes since trip start") +
  scale_fill_viridis(begin = 1, end = 0, name = "Speed (mph)") +#(name = "Speed (mph)", high = "#132B43", low = "#56B1F7") +
  #theme_bw() +
  theme(text = element_text(size=12), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), legend.position = c(.75, .25))

tripplot1 <- roadtype2 + ggtitle("A. Road type") +
  speed+ ggtitle("B. Speed")
tripplot1[[2]] = tripplot1[[2]] + theme(axis.text.y = element_blank(),
                                        axis.ticks.y = element_blank(),
                                        axis.title.y = element_blank() )

tripplot1

```


## Figure S`r k`. Variation in road density for highways and local roads at 500 m resolution (in standard deviations (SD)) and over each observed trip (N=69).  

```{r}
k <- k + 1
```


```{r}
rcomm1 <- tripplot[[2]]
```

```{r}
srness1 <- dplyr::select(rcomm1, ID, id2, timemin, contains("srness")) %>%
  pivot_longer(-c(ID, id2, timemin)) %>%
  dplyr::filter(name %in% c("srness500loc", "srness500hw")) %>%
  mutate(name = case_when(name == "srness500loc" ~ "Road density: Local",
                       
                          name == "srness500hw" ~ "Road density: Highway")) 



srness <- ggplot(srness1, aes(x = timemin, y = id2, fill = value)) +
  geom_tile() + ylab("Trip") + xlab("Minutes since trip start") +
 scale_fill_gradient2(name = "Road density", low = "#dedc59", high = "#5c3a85", mid = "white") +
  facet_wrap(~name) +
  theme(text = element_text(size=12),axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        legend.position  = c(.9, .3))  + theme( axis.text.y = element_blank(), axis.ticks.y = element_blank())

srness

```




## Figure S`r k`. Distributions of daily ambient PM$_{2.5}$ ($\mu$g/m$^3$) and hourly PM$_{2.5}$ ($\mu$g/m$^3$) for N=45 person-days of observation.

```{r}
k <- k + 1
```

```{r}
rcomm1 <- boxplot1[[2]]
ns <- boxplot1[[3]]

dates <- dplyr::select(rcomm1, date, ID) %>% unique() %>% rename(date_local = date)

# get data
va24 <- dailypm[[2]]
va24 <- full_join(va24, dates)


obsdiff0 <- fighourly[[2]]
obsdiff <- left_join(dates, obsdiff0)
alps <- fighourly[[3]] - 0.2
cols <- fighourly[[4]]


# add in median daily PM for hourly
va241 <- dplyr::select(va24, daily) %>% 
  mutate(type = "daily") 
sva <- summarize(va241, med = median(daily))
obsdiff1 <- dplyr::select(obsdiff, time_local, obsdiff) %>%
  mutate(type = "hourly", obsdiff = obsdiff + sva$med,
         type2 = paste0("Hour:", time_local)) %>%
  rename(pm = obsdiff) %>% dplyr::select(-time_local)
va241 <- dplyr::select(va24, daily) %>% 
  mutate(type = "Daily", type2 = "Daily") %>%
  rename(pm = daily)
vals <- c("Daily", paste0("Hour:", seq(0, 23)))
dhour <- bind_rows(obsdiff1, va241) %>%
  mutate(type2 = factor(type2, levels = vals))
```



```{r}
fighourly <- ggplot() +
  geom_boxplot(data = dhour, aes(x = factor(type2), y = pm, colour = type)) +
  geom_vline(xintercept = 1.5, colour = "grey20", linewidth = 1) +
  ylab(expression(paste("PM"[2.5], " (", mu, "g/m"^3, ")"))) +#, #"\nrelative to daily average"))) +
  xlab("") +
  scale_color_manual(values =c(1, cols[1]), guide = "none") +
  theme_bw() +
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
fighourly
```






## Figure S`r k`. Wind rose plot showing direction of the fastest wind (degrees) sustained over five minutes across N=45 person-days of observation.

```{r}
k <- k + 1
```

```{r}
plot.windrose(winddir1)

```










## Figure S`r k`. Associations of road density with in-vehicle PM$_{2.5}$ concentrations from linear mixed models reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of road density (Q2, Q3, Q4) to the lowest quartile (Q1) for highways and local roads at 500 m resolution. Results include main (unadjusted) models as well as models that separately adjusted for meteorology, ambient PM$_{2.5}$, rush hour trips, and all road features.

```{r}
k <- k + 1

```


```{r}
q500m
```





## Figure S`r k`. Associations of road type with in-vehicle PM$_{2.5}$ concentrations from linear mixed models, reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing travelling on highways, local connecting roads, and ramps/tunnels to local roads.  Results include main (unadjusted) models as well as models that separately adjusted for meteorology, ambient PM$_{2.5}$, and rush hour trips, and all road features.

```{r}
k <- k + 1
```


```{r}
fig5

```



## Figure S`r k`. Associations of speed with in-vehicle PM$_{2.5}$ concentrations from linear mixed models reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of speed (Q2, Q3, Q4) to the lowest quartile (Q1). Results include main (unadjusted) models as well as models that separately adjusted for meteorology, ambient PM$_{2.5}$, rush hour trips, and all road features.


```{r}
k <- k + 1

```


```{r}
qmph$fig
```


## Figure S`r k`. Associations of of road density with in-vehicle PM$_{2.5}$ concentrations from linear mixed models for main models (lag 0) and lags up to 10 minutes, reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of road density (Q2, Q3, Q4) to the lowest quartile (Q1) for highways and local roads at 500 m resolution.


```{r}
k <- k + 1
```

```{r}
fh <- 6
```

```{r, fig.height = fh}
roadlag500hw
```





## Figure S`r k`. Associations of road type with in-vehicle PM$_{2.5}$ concentrations from linear mixed models for main models (lag 0) and lags up to 10 minutes, reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing travelling on highways, ramps/tunnels, and local connecting roads to local roads.  


```{r}
k <- k + 1
```

```{r, fig.height = 7}
figs7

```



## Figure S`r k`. Associations of speed with in-vehicle PM$_{2.5}$ concentrations from linear mixed models for main models (lag 0) and lags up to 10 minutes, reported as the change in log $\mu$g/m$^3$ (95% confidence intervals) comparing each quartile of speed (Q2, Q3, Q4) to the lowest quartile (Q1).


```{r}
k <- k + 1
```

```{r, fig.height = fh}
mphlag

```


