---
title: 'Variability in PM2.5 exposures during daily commutes'
subtitle: "Miscellany"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = F)
```

```{r}
# libraries
library(tidyverse)
library(knitr)
library(here)
library(maps)
library(ggthemes)
library(DescTools)
library(RColorBrewer)
library(forcats)
library(lubridate)
library(patchwork)

# from roadiness-eda.Rmd
load(here("figures/eda-tabfig.RData"))
# from figure2.R
load(here("figures/fighourly.RData"))


# supplement from roadiness-eda.Rmd
load(here("results/supp-plots.RData"))
```


```{r}
rcomm1 <- boxplot1[[2]]
ns <- boxplot1[[3]]

dates <- select(rcomm1, date) %>% unique() %>% rename(date_local = date)
obsdiff <- fighourly[[2]]
obsdiff <- left_join(dates, obsdiff)

datecom <- dplyr::select(rcomm1, date, ID) %>% unique()

# plot of select
unid <- unique(rcomm1$id2)
# number of unique commutes
length(unid)
```

# Number of observations

- Number of obs `r nrow(rcomm1)`
- Number of trips `r length(unid)`
- Number of days `r length(unique(rcomm1$date))`
- Number of commuter-days `r nrow(datecom)`
- Number of commuters `r length(unique(datecom$ID))`

# Demographics

See roadiness-eda.Rmd

# Number of observations by season

```{r}
# number/prop by season

select(obsdiff, date_local, cold) %>% unique() %>% count(cold) %>%
  mutate(prop = n / sum(n)) %>%
  kable(digits = 2)

```


# PM extreme

Remove 2 observations with PM >= 250

```{r}
load(here("data/tripdata.RData"))
load(here("data/pm-cleaned.RData"))
pmlarge <- dplyr::filter(pm1b, PM >= 250)

tripdata1 <- dplyr::select(tripdata, ID, datetime, Latitude, Longitude) %>%
  mutate(rdatetime = round_date(datetime, "min"), hasll = 1) %>%
  na.omit() %>%
  dplyr::select(ID, rdatetime, hasll) %>%
  unique()
pmlarge1 <- left_join(pmlarge, tripdata1, by = c("ID", "rdatetime")) %>%
  na.omit()

kable(pmlarge1)

```



# commutes > 60


```{r}
load(here("data/rcomm2.RData"))
filter(rcomm2, obs > 60) %>% dplyr::select(ID, date, cum1, obs) %>%
  unique()

```

# Weather station

```{r}
load(here("data/weather-raw.RData"))
load(here("data/stations.RData"))

weather <- na.omit(weather)
uns <- unique(weather$id)

stations1 <- filter(stations, id %in% uns) %>%
  dplyr::select(id, latitude, longitude) %>% unique()

maps::map("county", xlim = c(-77.7, -77), ylim = c(38.6, 39.1))
points(stations1$longitude, stations1$latitude, col = "red")
```
# Monitor type?

```{r}
load(here("data/va_monitor_data.RData"))
with(va_monitor_data, table(method, AQS_site_ID))
with(va_monitor_data, table(method, sample_duration))
```
# make and model

```{r}
load("/Users/jenna/Documents/gestdc-data/gestdc-merged.RData")
load(here("figures/eda-tabfig.RData"))
rcomm1 <- boxplot1[[2]]
ids <- rcomm1$ID %>% unique() %>% tolower()
types <- c("small", "small", "crossover", "small", "crossover", "crossover")
dplyr::select(dat, id, modelcar, makecar) %>%
  dplyr::filter(modelcar != "" | makecar != "",
                id %in% ids) %>% mutate(type = types) %>%kable()


```


# Rush hour

```{r}
dplyr::select(rcommLM, ID, id3, rushmorn, rusheven) %>%
  mutate(none = ifelse(rushmorn == 0 & rusheven == 0, 1, 0),
         both =ifelse(rushmorn == 1 & rusheven == 1, 1, 0)) %>%
  #group_by(ID) %>%
  summarize(morn = mean(rushmorn), even = mean(rusheven),
            none = mean(none), both = mean(both))

```


# Roadiness and speed

```{r}
dplyr::select(rcommLM, ID, id3, mph, srness) %>%
  #group_by(ID) %>%
  summarize(speed = mean(mph), speedsd = sd(mph),
            msrness = mean(srness), srnesssd = sd(srness))

```


