---
title: 'Regression sensivity: quartiles'
author: "Jenna Krall"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
# libraries
library(tidyverse)
library(knitr)
library(splines)
library(here)
library(lme4)
library(broom.mixed)
library(ggthemes)
library(nlme)
library(forcats)
library(gtools)
library(patchwork)
library(RColorBrewer)
# library(lmeInfo)

source(here("functions/model-funs.R"))

source(here("functions/model-funs-quart.R"))


# load anonymous data
source(here("rcode/anon-data.R"))
rcommQ <- read.csv(here("data/roadden.csv"))
rcommQ <- rename(rcommQ, id3 = commid) %>%
  mutate(lPM = log(PM + 0.01))
rcommQ0 <- rcommQ

# load iqrs
load(here("data/iqrs.RData"))

```




# Test 1

Works when correlation is disabled, for all results

# Test 2

Works with sorted random ID, sorted random date, full correlation for all results

# Test 3 

- try unsorted, but still length 3: DOES NOT WORK
- Sorted date, not sorted ID: Does not work 
- Sorted ID, not sorted date: Does not work 
- using corAR1() instead of initializing does same thing
- using update instead of initializing does the same thing

# Test 4

Some issue with the observation order in corrStruc

- fine if use paste(ID, id3) only with update vs. initialize

```{r}
eval1 <- T
```

```{r, eval = eval1}
load(here("data/rcommQ.RData"))
rcommQ <- mutate(rcommQ, id3 = paste0(ID, id3))%>%    arrange(desc(ID), desc(id3), timemin)
print(head(unique(rcommQ$ID)))
print(head(unique(rcommQ$id3)))

lme1 <- lmequart(rcommQ, "qsrness1hw", "qmph")
lme1$t1[1 : 3, c(3, 4, 8 : 10)] 
t1 <- lme1$Main$t1[1 : 3, c(3, 4, 8 : 10)]
t1 %>% mutate(estimate = round(estimate, 2))
```


```{r, eval = eval1}
rcommQ <- rcommQ0
# rcommQ <- mutate(rcommQ, id3 = paste0(ID, id3))%>%    arrange(desc(ID), desc(id3), timemin)

print(head(unique(rcommQ$ID)))
print(head(unique(rcommQ$id3)))

lme1 <- lmequart(rcommQ, "qsrness1hw", "qmph")
# lme1$t1[1 : 3, c(3, 4, 8 : 10)] 

t1 <- lme1$Main$t1[1 : 3, c(3, 4, 8 : 10)]
t1 %>% mutate(estimate = round(estimate, 2))
```



```{r}

```

# Full


# Roadiness


## 1km res


```{r}
rcommQ <- rcommQ0
lme1 <- lmequart(rcommQ, "qsrness1loc", "qmph")
p1 <- plot_cat_sensQ(lme1, "qsrness1loc")


lme1 <- lmequart(rcommQ, "qsrness1hw", "qmph")
p2 <- plot_cat_sensQ(lme1, "qsrness1hw")

p1tab <- p1$tab %>% mutate(type = "Road density: Local")
p2tab <- p2$tab %>% mutate(type = "Road density: Highway")

pall <- full_join(p1tab, p2tab)
q1tab <- pall
q1km <- plot2(pall)
q1km

```



## 500 m res

```{r}
lme1 <- lmequart(rcommQ, "qsrness500loc", "qmph")
p1 <- plot_cat_sensQ(lme1, "qsrness500loc")


lme1 <- lmequart(rcommQ, "qsrness500hw", "qmph")
p2 <- plot_cat_sensQ(lme1, "qsrness500hw")

p1tab <- p1$tab %>% mutate(type = "Road density: Local")
p2tab <- p2$tab %>% mutate(type = "Road density: Highway")

pall <- full_join(p1tab, p2tab)
q500tab <- pall

q500m <- plot2(pall)
q500m
```


### Table of results

```{r}
q500tab1 <- dplyr::filter(q500tab, name == "Main") %>%
  mutate(Resolution = "500 m")
q1tab1 <- dplyr::filter(q1tab, name == "Main") %>%
  mutate(Resolution = "1 km")
r1 <- 2
qsrnesstab <- full_join(q500tab1, q1tab1) %>%
  dplyr::select(term, estimate, conf.low, conf.high, type, Resolution) %>%
  rename(Road = type) %>%
  mutate(estimate = round(estimate, r1),
         conf.low = round(conf.low, r1),
         conf.high = round(conf.high, r1),
         estCI = paste0(estimate, " (", conf.low, ", ", conf.high, ")")) %>%
  dplyr::select(term, estCI, Road, Resolution) %>%
  pivot_wider(names_from = "term", values_from = "estCI") 

qsrnesstab %>% kable()
```




# Old data


## 1km res


```{r}
load(here("data/rcommQ.RData"))
lme1 <- lmequart(rcommQ, "qsrness1loc", "qmph")
p1 <- plot_cat_sensQ(lme1, "qsrness1loc")


lme1 <- lmequart(rcommQ, "qsrness1hw", "qmph")
p2 <- plot_cat_sensQ(lme1, "qsrness1hw")

p1tab <- p1$tab %>% mutate(type = "Road density: Local")
p2tab <- p2$tab %>% mutate(type = "Road density: Highway")

pall <- full_join(p1tab, p2tab)
q1tab <- pall
q1km <- plot2(pall)

```



## 500 m res

```{r}
lme1 <- lmequart(rcommQ, "qsrness500loc", "qmph")
p1 <- plot_cat_sensQ(lme1, "qsrness500loc")


lme1 <- lmequart(rcommQ, "qsrness500hw", "qmph")
p2 <- plot_cat_sensQ(lme1, "qsrness500hw")

p1tab <- p1$tab %>% mutate(type = "Road density: Local")
p2tab <- p2$tab %>% mutate(type = "Road density: Highway")

pall <- full_join(p1tab, p2tab)
q500tab <- pall

q500m <- plot2(pall)

```


### Table of results

```{r}
q500tab1 <- dplyr::filter(q500tab, name == "Main") %>%
  mutate(Resolution = "500 m")
q1tab1 <- dplyr::filter(q1tab, name == "Main") %>%
  mutate(Resolution = "1 km")
r1 <- 2
qsrnesstab <- full_join(q500tab1, q1tab1) %>%
  dplyr::select(term, estimate, conf.low, conf.high, type, Resolution) %>%
  rename(Road = type) %>%
  mutate(estimate = round(estimate, r1),
         conf.low = round(conf.low, r1),
         conf.high = round(conf.high, r1),
         estCI = paste0(estimate, " (", conf.low, ", ", conf.high, ")")) %>%
  dplyr::select(term, estCI, Road, Resolution) %>%
  pivot_wider(names_from = "term", values_from = "estCI") 

qsrnesstab %>% kable()
```















# Speed

```{r, eval = F}
lme1 <- lmequart(rcommQ, "qmph", "qsrness1hw")
qmph <- plot_cat_sensQ(lme1, "qmph")
qmph$fig
```



